Improving phase-matching of FWM with temperature

### 2015-09-24 latex/davide_bazzanella.tex
stampa

@@

\subsection{Partially degenerate Four-Wave Mixing}
Four-Wave Mixing (FWM) is a nonlinear optical effect due to the susceptibility of silicon.
Silicon is an isotropic material and therefore second order susceptibility $\chi^{(2)}$ (almost) vanishes.
Third order susceptibility $\chi^{(3)}$, on the other hand, is non-zero and is responsible among the other effects for FWM, which is indeed a third order phenomenon.

@@

From the definition of $\phi_-$ and $\phi_+$ we can gather that there are two types of FWM.
The term containing $\phi_-$ corresponds to the case in which two photons at frequencies $\omega_1$ and $\omega_2$ are annihilated, while two photons at frequencies $\omega_3$ and $\omega_4$ are created simultaneously.
On the other hand, the term containing $\phi_+$ corresponds to the case in which three photons at frequencies $\omega_1$, $\omega_2$ and $\omega_3$ are annihilated, while one photon at frequency $\omega_4$ is created.

@@

We can also define a length scale, known as the coherence length, using $L_{coh} = 2\pi/|\Delta\kappa|$, where $\Delta\kappa$ is the maximum value of the effective phase mismatch that can be tolerated.
Moreover significant FWM occurs if $L_{sam}<L_{coh}$ where $L_{sam}$ is the sample length.
This definition allow us to make a further assumption, that is $\kappa = \Delta k + \gamma(P_1 + P_2) \approx \Delta k$.
Considering a peak power $P_1+P_2 = \SI{1}{\W}$ of the pump and a sample length $L_{sam} = \SI{1}{\cm}$ we obtain
\begin{equation*}
\left\{
 	\begin{aligned}
		\gamma(P_1 + P_2) &= \frac{n_2\omega}{cA_{eff}}P_0 = \frac{\SI{4.5e-18}{\m^2\per\W}}{c\SI{e-12}{\m^2}}\frac{2\pi c}{\SI{1.55e-6}{\m}}\SI{1}{\W} \approx \SI{20}{\per\m} \approx \SI{e1}{\per\m}\\
		|\Delta k |&= \frac{2\pi}{L_{sam}} = \frac{2\pi}{\SI{1}{\cm}} \approx \SI{600}{\per\m} \approx \num{e2} - \SI{e3}{\per\m}
	\end{aligned}
\right.
\label{eq_gamma_comparison}
\end{equation*}

@@

Therefore, in order to be efficient, the parametric process needs to verify the conservation of both energy and momentum, as given in Eq. (\ref{eq_f&pmc}).

@@

\begin{table}[ht]
\centering
  \begin{tabular}{lcccc}  
    \toprule
		& \multicolumn{3}{c}{Properties} \\
    \cmidrule(r){2-5}
    	\multirow{2}{*}{Materials}    & $\rho$ & $\kappa_{TH}$ & $C_P$ & TOC\\
  		& [\si{\kg\m^3}] & [\si{\W\per(\m\K)}] & [\si{\J\per(\kg\K)}] & [\si{\per\K}] \\
    \midrule
    Silicon (Si)			& 2329	& 130	& 700	& \num{1.87e-4}\\
    Silica (SiO$_2$)		& 2203	& 1.38	& 703	& 0\\
    \bottomrule
  \end{tabular}
  \label{tab_materials}
  \caption{Thermal properties of silicon and silica glass. \cite{n_db} and }
\end{table}

Moreover, in order to be influenced by the different temperatures, we had to add to the definition of refractive index of the silicon a term dependent on the temperature.
For our purpose this dependence can easily be considered linear:
\begin{equation}
n = n(\lambda)+\mathrm{TOC}\cdot (T-T_{amb})
\end{equation}
where $\mathrm{TOC}|_{\mathrm{Si}} = \SI{1.87e-4}{\per\K}$ is called Termo-Optic Coefficient. The coefficient of silica glass is negligible and therefore was considered zero $\mathrm{TOC}|_{\mathrm{SiO}_2} \approx 0$.

@@

@@

@@

This algorithm was set to find the eight modes whose effective index $n_{eff}$ was closest to $n_{eff}=4$ in absolute value.
\subsubsection{Data output}
We obtained the field distribution and a series of parameters as our data output for each parametric configuration.
In particular the parameters were the effective index $n_{eff}$, the effective area $A_{eff}$ and the nonlinear parameter $\gamma$, for each mode and for each wavelength and temperature.

\subsection{Data classification}
The next step was to classify these data, depending on the type (TE or TM) and order (1,2,3...) of the modes.
We achieved this by analyzing a sample taken at $y=0$ of the field distributions.
The ratio between the maxima values of the x-component and y-component of the electric field allowed to distinguish the mode type: TE or TM.
Those fields which had the maximum component in the y direction were categorized as TE modes while those which had the maximum component in the x direction as TM modes.
The number of maxima of the modulus of the field showed us the order of the mode.

We obtained then a well organized data matrix filled with the value of effective index $n_{eff}$, effective area $A_{eff}$ and nonlinear parameter $\gamma$ for each combination of parameters.

\subsection{Data processing}
We focused on the information given by the effective index.

@@

Specifically, we used a polynomial expression of fifth order in wavelength and second order in temperature ``\texttt{poly52}".

@@

We replaced its value in the conservation of momentum formula, giving us the phase mismatch as a function of signal wavelength and temperature of the heater only, for each set of orders.
\begin{align*}
	\Delta k &= 2\beta_P(\omega_P)-\beta_S(\omega_S)-\beta_I(2\omega_P-\omega_S)\\
			 &= 2\pi\left[ 2\frac{n_{eff}^P(\lambda_P)}{\lambda_P}
			 - \frac{n_{eff}^S(\lambda_S)}{\lambda_S}
			 - \frac{n_{eff}^I(\lambda_I)}{\lambda_I}\right]
\end{align*}

@@

Another important parameter is the width of the peak: to be consistent with the seletion criteria used in (\ref{eq_lcoh_crit}), we defined the bandwidth as the peak width at $L_{coh} = 1$.

We obtained graphs representing the position and the bandwidth of peaks vs temperature of each solution.
Figure (\ref{fig_example_position}) gives an example of peak position as a function of temperature $T_H$.
\begin{figure}[!hb]
	\centering
	\includegraphics[width=.8\textwidth]{ex_pos.png}
	\caption{Example of peak position vs $T_H$.}
	\label{fig_example_position}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=.75\textwidth]{variationTEsym.png}
	\caption{Percentage variation of $n_{eff}$ for TE modes at \SI{1.55}{\um} from its value at ambient temperature in the symmetric configuration.}
	\label{fig_variation}
\end{figure}

@@

In Figure (\ref{fig_variation}) is represented the change of effective index, for wavelength $\lambda = \SI{1.55}{\um}$, from ambient temperature to \SI{400}{\K} over it in ratio to its value at $T_{amb}=\SI{293.15}{\K}$.
This curve suggest us that higher orders are more influenced by the difference in temperature than the low orders.
We can therefore hypothesize that configuration with a high difference between their orders could be controlled better.
In this view, as next step in enhancing the influence of temperature on effective index, different core geometries should be studied.
%, in order to maximise the effect of the overlap integrals between the optical mode and the thermal profile.
This could greatly condition the wave propagation, and therefore the thermal tuning, but it is also a difficult problem with many degrees of freedom.

@@

\bibitem{agrawal} Agrawal, Govind. “Nonlinear Fiber Optics", 5th edition. Academic Press, 2013.
\bibitem{saleh} Saleh, Bahaa EA, and Malvin Carl Teich. “Fundamentals of photonics", 2nd edition. Wiley, 2007
\bibitem{n_db} http://refractiveindex.info/
\bibitem{spectrum} Lau, Ryan KW, et al. “Continuous-wave mid-infrared frequency conversion in silicon nanowaveguides." Optics letters 36.7 (2011): 1263-1265.
\bibitem{toc}	Frey, Bradley J., Douglas B. Leviton, and Timothy J. Madison. “Themperature-dependent refractive index of silicon and germanium." SPIE Astronomical Telescopes+ Instrumentation. International Society for Optics and Photonics, 2006.

>>>>

\vspace{20pt}
\noindent\textbf{Keywords}: photonics, SOI, multimode waveguide, nonlinear, FWM, thermo-optic coefficient, thermal tuning

@@

\subsection{Partially degenerate Four-Wave Mixing}
Four-Wave Mixing (FWM) is a nonlinear optical effect due to the susceptibility of silicon.
Silicon is an isotropic material and therefore second order susceptibility $\chi^{(2)}$ vanishes.
Third order susceptibility $\chi^{(3)}$, on the other hand, is non-zero and is responsible among the other effects for FWM, which is indeed a third order phenomenon.

@@

From the definition of $\phi_+$ and $\phi_-$ we can gather that there are two types of FWM.
The term containing $\phi_+$ corresponds to the case in which three photons at frequencies $\omega_1$, $\omega_2$ and $\omega_3$ are annihilated, while one photon at frequency $\omega_4$ is created.
On the other hand, the term containing $\phi_-$ corresponds to the case in which two photons at frequencies $\omega_1$ and $\omega_2$ are annihilated, while two photons at frequencies $\omega_3$ and $\omega_4$ are created simultaneously.

@@

We can also define a length scale, known as the coherence length, using $L_{coh} = 2\pi/|\Delta\kappa|$, where $\Delta\kappa$ is the maximum value of the effective phase mismatch that can be tolerated.
Moreover significant FWM occurs if $L_{sam}<L_{coh}$ where $L_{sam}$ is the sample length.
This definition allow us to make a further assumption, that is $\kappa = \Delta k + \gamma(P_1 + P_2) \approx \Delta k$.
Considering a peak power $P_1+P_2 = \SI{1}{\W}$ of the pump and a sample length $L_{sam} = \SI{1}{\cm}$ we obtain
\begin{equation}
\left\{
 	\begin{aligned}
		\gamma(P_1 + P_2) &= \frac{n_2\omega}{cA_{eff}}P_0 = \frac{\SI{4.5e-18}{\m^2\per\W}}{c\SI{e-12}{\m^2}}\frac{2\pi c}{\SI{1.55e-6}{\m}}\SI{1}{\W} \approx \SI{20}{\per\m} \approx \SI{e1}{\per\m}\\
		|\Delta k |&= \frac{2\pi}{L_{sam}} = \frac{2\pi}{\SI{1}{\cm}} \approx \SI{600}{\per\m} \approx \num{e2} - \SI{e3}{\per\m}
	\end{aligned}
\right.
\label{eq_gamma_comparison}
\end{equation}

@@

Therefore, in order to be efficient, the parametric process needs to verify the conservation of both energy and momentum, as given in Eq. (\ref{eq_f&pmc}):

\begin{equation*}
\left\{
 	\begin{aligned}
		\omega_I &= 2\omega_P - \omega_S \\
		\Delta k &= 2\beta_P - \beta_S - \beta_I = 0
	\end{aligned}
\right.
\end{equation*}

@@

\begin{table}[ht]
\centering
  \begin{tabular}{lcccc}  
    \toprule
		& \multicolumn{3}{c}{Properties} \\
    \cmidrule(r){2-5}
    	\multirow{2}{*}{Materials}    & $\rho$ & $\kappa_{TH}$ & $C_P$ & TOC\\
  		& [\si{\kg/\m^3}] & [\si{\W\per(\m\K)}] & [\si{\J\per(\kg\K)}] & [\si{\per\K}] \\
    \midrule
    Silicon (Si)			& 2329	& 130	& 700	& \num{1.87e-4}\\
    Silica (SiO$_2$)		& 2203	& 1.38	& 703	& 0\\
    \bottomrule
  \end{tabular}
  \label{tab_materials}
  \caption{Thermal properties of silicon and silica glass. \cite{n_db} and \cite{toc}}
\end{table}

Moreover, in order to be influenced by the different temperatures, we had to add to the definition of refractive index of the silicon a term dependent on the temperature.
For our purpose this dependence can easily be considered linear:
\begin{equation}
n = n(\lambda)+\mathrm{TOC}\cdot (T(x,y)-T_{amb})
\end{equation}
where $\mathrm{TOC}|_{\mathrm{Si}} = \SI{1.87e-4}{\per\K}$ is called Termo-Optic Coefficient. The coefficient of silica glass is negligible and therefore was considered zero $\mathrm{TOC}|_{\mathrm{SiO}_2} \approx 0$.

@@

\begin{figure}[!hb]
	\centering
	\includegraphics[width=1\textwidth]{thermal.pdf}
	\caption{Contour graph of a stationary solution for $T_H = T_{amb} + \SI{400}{\K}$}
	\label{fig_ex_thermal}
\end{figure}

@@

An example of solution is given in Figure (\ref{fig_ex_thermal}).

@@

This algorithm was set to find the fifteen modes whose effective index $n_{eff}$ was closest to $n_{eff}=4$ in absolute value.
\subsubsection{Data output}
We obtained the field distribution and a series of parameters as our data output for each parametric configuration.
In particular the parameters were the effective index $n_{eff}$ and the effective area $A_{eff}$, for each mode and for each wavelength and temperature.

\subsection{Data classification}
The next step was to classify these data, depending on the type (TE or TM) and order (1,2,3...) of the modes.
We achieved this by analyzing a sample taken at $y=0$ of the field distributions.
The ratio between the maxima values of the x-component and y-component of the electric field allowed to distinguish the mode type: TE or TM.
Those fields which had the maximum component in the x direction were categorized as TE modes while those which had the maximum component in the y direction as TM modes.
The number of maxima of the modulus of the field showed us the order of the mode.

We obtained then a well organized data matrix filled with the value of effective index $n_{eff}$ and effective area $A_{eff}$ for each combination of parameters.

\subsection{Data processing}
We used the $A_{eff}$ data to evaluate the nonlinear parameter $\gamma$ and verify Eq. (\ref{eq_gamma_comparison}).
We focused then on the information given by the effective index.

@@

Specifically, we used a polynomial expression of fifth order in wavelength and second order in temperature ``\texttt{poly52}", which is defined as:
\begin{align*}
	\texttt{poly52} = \,\,
		&p00 
		+ p10x + p01y
		+ p20x^2 + p11xy + p02y^2
		+ p30x^3 + p21x^2y + p12xy^2 +\\
+\,\,	&p40x^4 + p31x^3y + p22x^2y^2 
		+ p50x^5 + p41x^4y + p32x^3y^2
\end{align*}
where $x=\lambda$ and $y=T_H$

@@

We replaced its value in the conservation of momentum formula, giving us the phase mismatch as a function of signal wavelength and temperature of the heater only, for each set of orders.
\begin{align*}
	\Delta k &= 2\beta_P(\omega_P)-\beta_S(\omega_S)-\beta_I(2\omega_P-\omega_S)\\
			 &= 2\pi\left[ 2\frac{n_{eff}^P(\lambda_P,T_H)}{\lambda_P}
			 - \frac{n_{eff}^S(\lambda_S,T_H)}{\lambda_S}
			 - \frac{n_{eff}^I(\lambda_I,T_H)}{\lambda_I}\right]
\end{align*}

@@

Another important parameter is the width of the peak: to be consistent with the seletion criteria used in (\ref{eq_lcoh_crit}), we defined the bandwidth as the peak width at
$$L_{coh}/L_{sam} = 1$$

We obtained graphs representing the position and the bandwidth of peaks vs temperature of each solution.
Figures (\ref{fig_example_position}) and (\ref{fig_example_bandwidth}) give an example of peak position and bandwidth as a functions of temperature $T_H$.
\begin{figure}[!ht]
	\centering
	\includegraphics[width=.75\textwidth]{ex_pos.png}
	\caption{Example of peak position vs $T_H$, zoomed in an area between \SI{1.6}{\um} and \SI{1.61}{\um}. The smoothness of the curve is correlated to the step of evaluation in wavelength ($\Delta \lambda = \SI{0.05}{\nm}$)}
	\label{fig_example_position}
\end{figure}

\begin{figure}[!hb]
	\centering
	\includegraphics[width=.75\textwidth]{ex_bw.png}
	\caption{Example of bandwidth vs $T_H$. The figure has a little ripple which is correlated to the step of evaluation in wavelength ($\Delta \lambda = \SI{0.05}{\nm}$).}
	\label{fig_example_bandwidth}
\end{figure}

@@

\begin{figure}[!ht]
	\centering
	\includegraphics[width=.75\textwidth]{variation.png}
	\caption{Percentage variation of $n_{eff}$ for TE modes at \SI{1.55}{\um} from its value at ambient temperature in the symmetric configuration.}
	\label{fig_variation}
\end{figure}

@@

In Figure (\ref{fig_variation}) is represented the change of effective index, for wavelength $\lambda = \SI{1.55}{\um}$, from ambient temperature to \SI{400}{\K} over it in ratio to its value at $T_{amb}=\SI{293.15}{\K}$.
This curve suggest us that low orders are more influenced by the difference in temperature than high orders.
We can therefore hypothesize that configuration with a high difference between their orders could be controlled better.
In this view, as next step in enhancing the influence of temperature on effective index, different core geometries should be studied.
%, in order to maximise the effect of the overlap integrals between the optical mode and the thermal profile.
This could greatly condition the wave propagation, and therefore the thermal tuning, but it is also a difficult problem with many degrees of freedom.

@@

\bibitem{agrawal} Agrawal, Govind. “Nonlinear Fiber Optics", 5th edition. Academic Press, 2013.
\bibitem{saleh} Saleh, Bahaa EA, and Malvin Carl Teich. “Fundamentals of photonics", 2nd edition. Wiley, 2007
\bibitem{spectrum} Lau, Ryan KW, et al. “Continuous-wave mid-infrared frequency conversion in silicon nanowaveguides." Optics letters 36.7 (2011): 1263-1265.
\bibitem{toc}	Frey, Bradley J., Douglas B. Leviton, and Timothy J. Madison. “Themperature-dependent refractive index of silicon and germanium." SPIE Astronomical Telescopes+ Instrumentation. International Society for Optics and Photonics, 2006.
\bibitem{n_db} http://refractiveindex.info/
